Supplementary Information

### 2022-02-23 writing/appendix_a.Rmd
Tidying and removing sections that are duplicated in paper.Rmd

@@

```{r, big-discrepacies, eval=FALSE}
drugs <- unique(differences$drug)

@@

```{r, gls-functions}
drugs <- unique(differences$drug)

#### Prep GLS data ####
glsData <- function(data){
  df <- 
    differences %>% 
    filter(drug == data) %>% 
    mutate(Month = as.numeric(as.factor(datename)),
           pct_change = 100*(prop - 1)) %>% 
    select(c(regional_unit, Month, pct_change))

  baseline <- subset(df, Month == 2, -Month)
  baseline <- upData(baseline, rename = c(pct_change = "feb_diff"), print = FALSE)
  followup <- subset(df, Month > 2)
  both <- merge(baseline, followup, by = c("regional_unit"))

  #### Join GP and GLS data ####
  both <- 
    both %>% 
    left_join(gp_2020, by = "regional_unit") %>% 
    mutate(gp_load = npat / ngp)

  dd <- datadist(both)
  options(datadist = "dd")

  return(both)
}

#### Redundancy analysis model ####
redun <- list()
for(i in drugs){
  redun[[i]] <- redun(~hboard + dispensing + Month + feb_diff + wimd2019 + 
                        PC1 + PC2 + PC3 + PC4 + PC5 + ngp + gp_load,
                      data = glsData(i), r2 = .8, type = "adjusted") %>% 
    .$Out
}

# No redundent variables

#### GLS Analysis function ####
bestGLS <- function(data, bootstrap = FALSE){
  cp <- list(corExp, corCompSymm, corLin, corGaus, corSpher)
  z <- vector("list", length(cp))

  ### Fit models
  for(k in 1:length(cp)){
    z[[k]] <- gls(pct_change ~ rcs(Month, 4:7) +
                    hboard + dispensing +
                    rcs(wimd2019, 4) + 
                    rcs(ngp, 4) + rcs(gp_load, 4) + 
                    rcs(PC1, 4) + rcs(PC2, 4) + rcs(PC3, 4) + 
                    rcs(PC4, 4) + rcs(PC5, 4) +
                    hboard:rcs(Month, 4:7) +
                    dispensing:rcs(Month, 4:7) +
                    rcs(wimd2019, 4):rcs(Month, 4:7) +
                    rcs(ngp, 4):rcs(Month, 4:7) +
                    rcs(gp_load, 4):rcs(Month, 4:7),
                  data = data,
                  correlation = cp[[k]](form = ~ Month | regional_unit))
  }

  arma_tab <- # Table for p and q values for corARMA()
    tibble(p = 1:3, 
           q = rep(0,3)) %>% 
    mutate(model = row_number() + length(cp))

  z <- c(z, map2(.x = arma_tab$p, .y = arma_tab$q, 
                 .f = ~ if(.x > 0 | .y >0)
                 { # p = 0 & q = 0 fails
                   tryCatch(
                     {gls(pct_change ~ rcs(Month, 4:7) +
                            hboard + dispensing +
                            rcs(wimd2019, 4) + 
                            rcs(ngp, 4) + rcs(gp_load, 4) + 
                            rcs(PC1, 4) + rcs(PC2, 4) + rcs(PC3, 4) + 
                            rcs(PC4, 4) + rcs(PC5, 4) +
                            hboard:rcs(Month, 4:7) +
                            dispensing:rcs(Month, 4:7) +
                            rcs(wimd2019, 4):rcs(Month, 4:7) +
                            rcs(ngp, 4):rcs(Month, 4:7) +
                            rcs(gp_load, 4):rcs(Month, 4:7),
                          data = data,
                          correlation = corARMA(form = ~ Month|regional_unit, 
                                                p=.x, q=.y))
                     },
                     error = function(cond){
                       message(paste("Model causing error: p = ", 
                                     .x, "q = ", .y))
                       message("Original error message:")
                       message(cond)
                       return(NULL)
                     },
                     warning = function(cond){
                       message(paste("Model causing warning: p = ", 
                                     .x, "q = ", .y))
                       message("Original warning message:")
                       message(cond)
                       return(NULL)
                     }
                   )
                 }))

  ### Select best model
  names(z) <- paste0("model", 1:length(z))
  x <- z %>% 
    map(summary)

  # Get metrics for each model
  aic <- map_dfr(x, pluck("AIC"), .id = "model") 
  bic <- map_dfr(x, pluck("BIC"), .id = "model")
  log_lik <- map_dfr(x, pluck("logLik"), .id = "model")

  # Create table
  mod_comparison <- 
    bind_cols(metric = c("aic", "bic", "log_lik"), 
              bind_rows(aic,bic,log_lik)) %>% 
    pivot_longer(-metric) %>% 
    pivot_wider(id_cols = name, values_from = value, names_from = metric) %>% 
    rename(model = name) %>% 
    mutate(model = as.numeric(str_remove_all(model, "model")))

  # Select best model
  best_model_aic <- 
    mod_comparison %>% 
    arrange(aic, bic, desc(log_lik)) %>% 
    head(1) %>% 
    .$model

  ### Fit best model
  if(bootstrap == TRUE){  
    if(best_model_aic <= 6){
      a <-  Gls(pct_change ~ rcs(Month, 4:7) +
                  hboard + dispensing +  
                  rcs(wimd2019, 4) + 
                  rcs(ngp, 4) + rcs(gp_load, 4) + 
                  rcs(PC1, 4) + rcs(PC2, 4) + rcs(PC3, 4) + 
                  rcs(PC4, 4) + rcs(PC5, 4) +
                  hboard:rcs(Month, 4:7) +
                  dispensing:rcs(Month, 4:7) +
                  rcs(wimd2019, 4):rcs(Month, 4:7) +
                  rcs(ngp, 4):rcs(Month, 4:7) +
                  rcs(gp_load, 4):rcs(Month, 4:7),
                data = data, B = 1000,
                correlation = 
                  cp[[best_model_aic]](form = ~ Month | regional_unit))
      }
    else{
      a <-  Gls(pct_change ~ rcs(Month, 4:7) +
                  hboard + dispensing +  
                  rcs(wimd2019, 4) + 
                  rcs(ngp, 4) + rcs(gp_load, 4) + 
                  rcs(PC1, 4) + rcs(PC2, 4) + rcs(PC3, 4) + 
                  rcs(PC4, 4) + rcs(PC5, 4) +
                  hboard:rcs(Month, 4:7) +
                  dispensing:rcs(Month, 4:7) +
                  rcs(wimd2019, 4):rcs(Month, 4:7) +
                  rcs(ngp, 4):rcs(Month, 4:7) +
                  rcs(gp_load, 4):rcs(Month, 4:7),
                data = data, B = 1000,
                correlation = 
                  corARMA(form = ~ Month | regional_unit, 
                          p = filter(arma_tab, model == best_model_aic)$p, 
                          q = filter(arma_tab, model == best_model_aic)$q))
      }
  }
  else{
    if(best_model_aic <= 6){
      a <- Gls(pct_change ~ rcs(Month, 4:7) +
                 hboard + dispensing +  
                 rcs(wimd2019, 4) + 
                 rcs(ngp, 4) + rcs(gp_load, 4) + 
                 rcs(PC1, 4) + rcs(PC2, 4) + rcs(PC3, 4) + 
                 rcs(PC4, 4) + rcs(PC5, 4) +
                 hboard:rcs(Month, 4:7) +
                 dispensing:rcs(Month, 4:7) +
                 rcs(wimd2019, 4):rcs(Month, 4:7) +
                 rcs(ngp, 4):rcs(Month, 4:7) +
                 rcs(gp_load, 4):rcs(Month, 4:7),
               data = data,
               correlation = 
                 cp[[best_model_aic]](form = ~ Month | regional_unit))
    }
    else{
      a <- Gls(pct_change ~ rcs(Month, 4:7) +
                  hboard + dispensing +  
                  rcs(wimd2019, 4) + 
                  rcs(ngp, 4) + rcs(gp_load, 4) + 
                  rcs(PC1, 4) + rcs(PC2, 4) + rcs(PC3, 4) + 
                  rcs(PC4, 4) + rcs(PC5, 4) +
                  hboard:rcs(Month, 4:7) +
                  dispensing:rcs(Month, 4:7) +
                  rcs(wimd2019, 4):rcs(Month, 4:7) +
                  rcs(ngp, 4):rcs(Month, 4:7) +
                  rcs(gp_load, 4):rcs(Month, 4:7),
               data = data,
               correlation = 
                 corARMA(form = ~ Month | regional_unit, 
                         p = filter(arma_tab, model == best_model_aic)$p, 
                         q = filter(arma_tab, model == best_model_aic)$q))
    }
  }
  return(a)
}

#### GLS Assumption Checking ####
glsAssump <- function(model, data){
  data$resid <- r <- resid(model); data$fitted <- fitted(model)
  y1 <- ylab("Residuals")
  p1 <- ggplot(data, aes(x = fitted, y = resid)) + geom_point(alpha = .1) + y1
  p2 <- ggplot(data, aes(x = feb_diff, y = resid)) + geom_point(alpha = .1) + y1
  p3 <- ggplot(data, aes(x = Month, y = resid)) + y1 + 
    stat_summary(fun.data = "mean_sdl", geom = "smooth")
  p4 <- ggplot(data, aes(sample = resid)) + stat_qq(alpha = .1) + 
    geom_abline(intercept = mean(r), slope = sd(r)) + y1

  gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
}

@@

  p1 <- Predict(model, Month, dispensing, hboard,
                PC1 = percentiles(data, PC1),
                conf.int = .95) %>% 
    predictPlot(., groups = PC1, "PC1")
  p2 <- Predict(model, Month, dispensing, hboard,
                PC2 = percentiles(data, PC2),
                conf.int = .95) %>% 
    predictPlot(., groups = PC2, "PC2")
  p3 <- Predict(model, Month, dispensing, hboard,
                PC3 = percentiles(data, PC3),
                conf.int = .95) %>% 
    predictPlot(., groups = PC3, "PC3")
  p4 <- Predict(model, Month, dispensing, hboard,
                PC4 = percentiles(data, PC4),
                conf.int = .95) %>% 
    predictPlot(., groups = PC4, "PC4")
  gridExtra::grid.arrange(p1,p2,p3,p4, ncol = 2, newpage = TRUE) # Effects 6
  cat("\n\n")

@@

```{r, gls-analysis-save, results="asis", fig.cap=gls_analysis_fig_cap, warning=FALSE, message=FALSE, eval = FALSE}
#### gls-analysis-save ####
drugs <- unique(differences$drug)

best_models <- list()
best_bs_models <- list()

dd <- NULL
for(i in drugs){
  gls_data <- glsData(i)
  dd <- datadist(gls_data)
  options(datadist = "dd")
  best_gls <- bestGLS(gls_data)
  best_bs_gls <- bestGLS(gls_data, TRUE)
  rm(dd)
  best_models[[i]] <-
    best_gls
  best_bs_models[[i]] <- 
    best_bs_gls
  options(datadist = NULL)
}

write_rds(best_models, "../data/gls/best_models.rds")
write_rds(best_bs_models, "../data/gls/best_bs_models.rds")
```

@@

```{r, gls-effects, results="asis", fig.cap=gls_analysis_fig_cap, dependson="gls-functions"}
#### gls-effects ####
drugs <- unique(differences$drug)

@@

```{r, gls-anova-summary-tab}
#### gls-anova-summary-tab ####
best_bs_models %>% 
  map(~anova(., tol = 1e-10)) %>% 
  map(~as_tibble(x = ., rownames = "Coef")) %>% 
  map(as.matrix) %>% 
  map(as_tibble) %>% 
  bind_rows(.id= "drug") %>% 
  mutate(across(`Chi-Square`:P, as.numeric)) %>% 
  select(-c(`Chi-Square`, d.f.)) %>% 
  filter(!str_detect(Coef, "Nonlinear") &
           !str_detect(Coef, "TOTAL") &
           !str_detect(Coef,"All Interactions"),
         !str_detect(Coef, c("f" %R% OPEN_PAREN %R% "A,B"))) %>%
  pivot_wider(names_from = drug, values_from = P) %>% 
  mutate(across(ace:warfarin, ~printnum(., zero = FALSE)),
         Coef = str_remove_all(Coef, OPEN_PAREN %R% "Factor" %R% PLUS %R% 
                                 "Higher Order Factors" %R% CLOSE_PAREN)) %>% 
  select(Coef, all_drugs, everything()) %>% 
  kable(booktabs = TRUE, linesep = "",
        caption = paste("Wald test p-values"))%>% 
  kable_styling(latex_options = c("striped", "HOLD_position", "repeat_header")) %>% 
  landscape()
```

>>>>

drugs <- unique(differences$drug)

@@

```{r, big-discrepacies, eval=FALSE}

@@

```{r, gls-functions}

#### Prep GLS data ####
glsData <- function(data){
  df <- 
    differences %>% 
    filter(drug == data) %>% 
    mutate(Month = as.numeric(as.factor(datename)),
           pct_change = 100*(prop - 1)) %>% 
    select(c(regional_unit, Month, pct_change)) %>% 
    filter(Month > 2) %>% 
    left_join(gp_2020, by = "regional_unit") %>% 
    mutate(gp_load = npat / ngp)

  dd <- datadist(df)
  options(datadist = "dd")

  return(df)
}

#### GLS Assumption Checking ####
glsAssump <- function(model, data){
  data$resid <- r <- resid(model); data$fitted <- fitted(model)
  y1 <- ylab("Residuals")
  p1 <- ggplot(data, aes(x = fitted, y = resid)) + geom_point(alpha = .1) + y1
  # p2 <- ggplot(data, aes(x = feb_diff, y = resid)) + geom_point(alpha = .1) + y1
  p3 <- ggplot(data, aes(x = Month, y = resid)) + y1 + 
    stat_summary(fun.data = "mean_sdl", geom = "smooth")
  p4 <- ggplot(data, aes(sample = resid)) + stat_qq(alpha = .1) + 
    geom_abline(intercept = mean(r), slope = sd(r)) + y1

  gridExtra::grid.arrange(p1, #p2, 
                          p3, p4, ncol = 2)
}

@@

  # p1 <- Predict(model, Month, dispensing, hboard,
  #               PC1 = percentiles(data, PC1),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC1, "PC1")
  # p2 <- Predict(model, Month, dispensing, hboard,
  #               PC2 = percentiles(data, PC2),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC2, "PC2")
  # p3 <- Predict(model, Month, dispensing, hboard,
  #               PC3 = percentiles(data, PC3),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC3, "PC3")
  # p4 <- Predict(model, Month, dispensing, hboard,
  #               PC4 = percentiles(data, PC4),
  #               conf.int = .95) %>% 
  #   predictPlot(., groups = PC4, "PC4")
  # gridExtra::grid.arrange(p1,p2,p3,p4, ncol = 2, newpage = TRUE) # Effects 6
  # cat("\n\n")

@@

@@

```{r, gls-effects, results="asis", fig.cap=gls_analysis_fig_cap, dependson="gls-functions"}
#### gls-effects ####

@@

```{r, gls-anova-summary-tab}
#### gls-anova-summary-tab ####
best_bs_models %>% 
  map(~anova(., tol = 1e-10)) %>% 
  map(~as_tibble(x = ., rownames = "Coef")) %>% 
  map(as.matrix) %>% 
  map(as_tibble) %>% 
  bind_rows(.id= "drug") %>% 
  mutate(across(`Chi-Square`:P, as.numeric)) %>% 
  select(-c(`Chi-Square`, d.f.)) %>% 
  filter(!str_detect(Coef, "Nonlinear") &
           !str_detect(Coef, "TOTAL") &
           !str_detect(Coef,"All Interactions"),
         !str_detect(Coef, c("f" %R% OPEN_PAREN %R% "A,B"))) %>%
  pivot_wider(names_from = drug, values_from = P) %>% 
  mutate(across(ace:warfarin, ~printnum(., zero = FALSE)),
         Coef = str_remove_all(Coef, OPEN_PAREN %R% "Factor" %R% PLUS %R% 
                                 "Higher Order Factors" %R% CLOSE_PAREN)) %>% 
  select(Coef, all_drugs, everything()) %>% 
  kable(booktabs = TRUE, linesep = "",
        caption = paste("Wald test p-values")) %>% 
  kable_styling(latex_options = c("striped", "HOLD_position", "repeat_header")) %>% 
  landscape()
```
